Mirror, mirror on the wall. Who's the best match of them all?

### 2023-11-24 PNAS/index.Rnw
Add figures and newlines
Not sure how to get the figure to be the right size... ugh
Updates

This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehlerFingerprintErrorRate2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
\svp{Very close non-matches are a direct consequence of a large number of comparisons; they will occur more frequently as databases increase in size and algorithms become more powerful.}
% In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.

In order to avoid embarassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
%
% \hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. } % This doesn't seem super-relevant and distracts from the continuity between the algorithm discussion above and the next paragraph

@@

Algorithms make a large number of comparisons \svp{that are not obvious to the user}
%out of a user's sight.
The cross-correlation function, for example, \svp{was} \citep{vorburgerApplicationsCrosscorrelationFunctions2011} one of the first quantitative measures \svp{proposed} in response to the NAS report \citep{nas2009}, and is used in \svp{many} pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
Finding the best \svp{alignment} requires implicitly sliding one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source \svp{ while keeping} track of \svp{the value of a} similarity measure.
\svp{This mirrors the forensic examination process, as examiners visually}
rotate or align two pieces of evidence under comparison microscopes.

%An explicit search, in constrast, occurs when we conduct database searches.
%These searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.

@@

A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts which can be compared to the cut surface of the wire. These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface. The blade cuts will then be compared to the wire under a comparison microscope; eventually, automatic comparison algorithms may also be validated for lab use. Each side of each blade cut will be compared to each side of the wire; depending on the tool design, there can be between 2 and 4 cutting surfaces in contact with the substrate.

@@

@@

\subsection*{A concrete example}
Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface; the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two. This is a minimal scenario - the wire will have striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B. The blade cuts of a sheet of aluminum that is 1.5 cm in width thus produce two striated edges corresponding to side A and side B which can be compared to other cuts to assess similarity.

>>>>

This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
\svp{Very close non-matches are a direct consequence of a large number of comparisons; they will occur more frequently as databases increase in size and algorithms become more powerful.}
% In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.

%
% \hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. } % This doesn't seem super-relevant and distracts from the continuity between the algorithm discussion above and the next paragraph

@@

Algorithms make a large number of comparisons \svp{that are not obvious to the user}
%out of a user's sight.
The cross-correlation function, for example, \svp{was} \citep{vorburgerApplicationsCrosscorrelationFunctions2011} one of the first quantitative measures \svp{proposed} in response to the NAS report \citep{nas2009}, and is used in \svp{many} pattern searching algorithms to find the best alignment between two sources and also evaluate the overall similarity.
Finding the best \svp{alignment} requires implicitly sliding one source across the whole length (for one-dimensional sources, such as striations) or area (for two dimensional sources, such as impression marks) of the other source \svp{ while keeping} track of \svp{the value of a} similarity measure.
\svp{This mirrors the forensic examination process, as examiners visually}
rotate or align two pieces of evidence under comparison microscopes.
In order to avoid embarassing false accusations and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.
%An explicit search, in constrast, occurs when we conduct database searches.
%These searches often return the $N$ best matches from the database, but this requires making comparisons to all of the database entries which meet the search criteria.

@@

A forensics examiner tasked with determining whether a wire in evidence was cut by a specific tool will create one or more blade cuts which can be compared to the cut surface of the wire.
These cuts are made in a sheet of material similar to the wire's composition, and may be performed at multiple angles, as the angle of the tool to the substrate can affect which striations are recorded on the substrate surface.
The blade cuts will then be compared to the wire under a comparison microscope in a process shown in \Cref{fig:wire-blade}, though eventually, automatic comparison algorithms may also be validated for lab use.
Each side of each blade cut will be compared to each side of the wire; depending on the tool design, there can be between 2 and 4 cutting surfaces in contact with the substrate.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{../fig/wire-blade-comparison.png}
\caption{A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match).}\label{fig:wire-blade}
\end{figure}

@@

\subsection*{A concrete example}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{../fig/wire_cutter_wires.png}
\caption{RJ45 Crimp tool with a 1.5cm razor blade used for cutting. 1mm and 2mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}
\end{figure}

Let us consider a wire-cutting tool with a 1.5 cm razor blade that meets a cast surface (one such tool is shown in \Cref{fig:tool-pic}); the wire is held against this rectangular cast surface as the blade is pushed into the wire, splitting it in two.
This is a minimal scenario - the wire will have striations from one side of the blade, while the blade itself has two cutting edges, which we will call side A and side B.
The blade cuts of a sheet of aluminum that is 1.5 cm in width thus produce two striated edges corresponding to side A and side B which can be compared to other cuts to assess similarity.
